Case for Support

### 2021-11-17 application/backgroundToTheResource.tex

To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour whilst recording, and often manipulating, neural activity. 
%
Powerful experimental designs require behavioural feedback (e.g.\ reward, lights, sounds, actuated
devices) or neural perturbation (e.g.\ using light-sensitive polarising
molecules) that depend on real-time behaviour or neural state.
%
It has also become increasingly apparent that the neural signals that
drive complex behaviour are manifest in a broad population of neurons rather
than in single cells. Current neural recording emphasises high-dimensional
signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
%
\textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}. The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining data from old ones.

>>>>

To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour while recording, and often manipulating, neural activity. 
%
Powerful experimental designs require behavioural feedback (e.g.\ reward, lights, sounds, actuated
devices) or neural perturbation (e.g.\ using light-sensitive polarising
molecules) that depend on real-time behaviour or neural state.
%
It has also become increasingly apparent that the neural signals that
drive complex behaviour are manifest in a broad population of neurons rather
than in single cells. Current neural recording emphasises high-dimensional
signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
%
\textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}. The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining data from old ones.

### 2021-11-17 application/detailsOfResource.tex

As it stands, Bonsai is designed to be extensible and is distributed with a package-management framework that controls installation of optional core packages as well as user-contributed extensions.
%
However extensions must be written in Bonsai's native C\# programming language.
%
This specificity restricts the range of potential contributors to the ecosystem.
%
It also hinders easy incorporation of published analytic tools, many of which are implemented within other interpreted language systems such as Python, R, or MATLAB.  
\begin{wrapfigure}{R}{0pt}
  \includegraphics[width=0.5\textwidth]{figures/extensions.pdf}
%   \input{figures/extensions}
  \caption{Proposed extensions to Bonsai}
  \label{fig:proposedBonsaiExtensions}
\end{wrapfigure}

@@

To avoid the need to re-implement a single data analysis algorithm to handle multiple different data types we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and machine-learning algorithms.
%
The planned abstraction layer is based on the successful \texttt{scikit-learn} Python
library \citep{buitinckEtAl13}.
%
By analogy to the interfaces provided by this library we will introduce three types of Bonsai operator: \texttt{Estimator} objects for building and fitting models, \texttt{Predictor} objects for making predictions and \texttt{Transformer} objects for converting data.
%
Interactions between these different object types will allow a single algorithm class (implemented by \texttt{Estimator} and \texttt{Predictor} operators) to act on many different data streams.
%
% An example closed-loop architecture allowing online control of a compu|ter cursor with neural activity is shown in Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
% graphical program for this system is given in Figure~\ref{fig:proposedBonsaiExtensions}c. The
% objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
% GPFAlatentsExtractor implement the \texttt{Transformer} interface (blue),
% and the object instantiating the TimeSeriesForestClassifier implements the
% \texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
% %
% While the parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter may
% be set manually by the user, the parameters of
% the GPFAlatentsExtractor can be learnt offline in an unsupervised way using an
% \texttt{Estimator} (red) in a second pipeline
% (Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
% Finally, the parameters of the TimeSeriesForestClassifier can also be learnt
% offline, but in a supervised way, using another \texttt{Estimator} in a third
% pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).

@@

\subsubsection{Video-based behavioral analysis}
\label{sec:videoBasedBehavioralAnalysis}

@@

    \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15}: extracts patterns
        of behaviour that repeat over time (i.e., syllables of behaviour) from
        video data. For instance, it parses (in an unsupervised way) the
        behaviour of a mouse in an arena into segments of time where a mouse
        was running, rearing and grooming. It uses a hidden Markov model and it
        is implemented in Python\footnote{Code for MoSeq can be requested from
        the Datta laboratory, as indicated at
        http://datta.hms.harvard.edu/research/behavioral-analysis/.}.
        After trained MoSeq can be used to detect behavioral syllables online.

@@

We propose to add functionality into Bonsai for the implementation of
brain-computer-interface (BCI) applications illustrated in
Figure~\ref{fig:proposedBonsaiExtensions}b. In
this implementation voltage recordings from the brain will be converted into
spikes fired by single neurons by a \texttt{SpikeDetector}
(Section~\ref{sec:spikeDetection}), or to local field potentials (LFP) by an
\texttt{LFPextractor} (Section~\ref{sec:lfpExtraction}). Next, neural activity
(spikes or LFPs) will be represented in a low-dimensional latent space by a
\texttt{LatentsCalculator}. Finally, a \texttt{Decoder} will extract the
intended behavior of the animal from this low-dimensional representation.

@@

We propose to assess the progress of the proposed project using three different
types of targets.
%
First, we will have targets associated with each release to the general public
of the software components described above. 
% in Sections~\ref{sec:infrastructure} and~\ref{sec:functionality}.
%
Second, we will monitor user activity in the Bonsai
forum\footnote{https://groups.google.com/g/bonsai-users},
and publications citing Bonsai that use th new machine larning functionality,
to estimate the number of Bonsai users using this functionality.
%
Third, we will count the number of machine learning methods that are integrated
into the Bonsai ecosystem by us and by other method developers.

>>>>

As it stands, Bonsai is designed to be extensible and is distributed with a package-management framework that controls installation of optional core packages as well as user-contributed extensions.
%
However extensions must be written in Bonsai's native C\# programming language.
%
This specificity restricts the range of potential contributors to the ecosystem.
%
It also hinders easy incorporation of published analytic tools, many of which are implemented within other interpreted language systems such as Python, R, or MATLAB.  
\begin{wrapfigure}{R}{0pt}
  \includegraphics[width=0.5\textwidth]{figures/extensions.pdf}%
%   \input{figures/extensions}
  \caption{Proposed extensions to Bonsai}
  \label{fig:proposedBonsaiExtensions}
\end{wrapfigure}

@@

To avoid the need to re-implement a single data analysis algorithm to handle multiple different data types we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and ML algorithms.
%
The planned abstraction layer is based on the successful \texttt{scikit-learn} Python
library \citep{buitinckEtAl13}.
%
By analogy to the interfaces provided by this library we will introduce three types of Bonsai operator: \texttt{Estimator} objects for building and fitting models, \texttt{Predictor} objects for making predictions and \texttt{Transformer} objects for converting data.
%
Interactions between these different object types will allow a single algorithm class (implemented by \texttt{Estimator} and \texttt{Predictor} operators) to act on many different data streams.
%
% An example closed-loop architecture allowing online control of a compu|ter cursor with neural activity is shown in Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
% graphical program for this system is given in Figure~\ref{fig:proposedBonsaiExtensions}c. The
% objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
% GPFAlatentsExtractor implement the \texttt{Transformer} interface (blue),
% and the object instantiating the TimeSeriesForestClassifier implements the
% \texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
% %
% While the parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter may
% be set manually by the user, the parameters of
% the GPFAlatentsExtractor can be learnt offline in an unsupervised way using an
% \texttt{Estimator} (red) in a second pipeline
% (Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
% Finally, the parameters of the TimeSeriesForestClassifier can also be learnt
% offline, but in a supervised way, using another \texttt{Estimator} in a third
% pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).

@@

\subsubsection{Video-based behavioural analysis}
\label{sec:videoBasedBehavioralAnalysis}

@@

    \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15}: extracts patterns
        of behaviour that repeat over time (i.e., syllables of behaviour) from
        video data. For instance, it parses (in an unsupervised way) the
        behaviour of a mouse in an arena into segments of time where a mouse
        was running, rearing and grooming. It uses a hidden Markov model and it
        is implemented in Python\footnote{Code for MoSeq can be requested from
        the Datta laboratory, as indicated at
        http://datta.hms.harvard.edu/research/behavioral-analysis/.}.
        After trained MoSeq can be used to detect behavioural syllables online.

@@

We propose to add functionality into Bonsai for the implementation of
brain-computer-interface (BCI) applications illustrated in
Figure~\ref{fig:proposedBonsaiExtensions}b. In
this implementation voltage recordings from the brain will be converted into
spikes fired by single neurons by a \texttt{SpikeDetector}
(Section~\ref{sec:spikeDetection}), or to local field potentials (LFP) by an
\texttt{LFPextractor} (Section~\ref{sec:lfpExtraction}). Next, neural activity
(spikes or LFPs) will be represented in a low-dimensional latent space by a
\texttt{LatentsCalculator}. Finally, a \texttt{Decoder} will extract the
intended behaviour of the animal from this low-dimensional representation.

@@

We propose to assess the progress of the proposed project using three different
types of targets.
%
First, we will have targets associated with each release to the general public
of the software components described above. 
% in Sections~\ref{sec:infrastructure} and~\ref{sec:functionality}.
%
Second, we will monitor user activity in the Bonsai
forum\footnote{https://groups.google.com/g/bonsai-users},
and publications citing Bonsai that use the new ML functionality,
to estimate the number of Bonsai users using this functionality.
%
Third, we will count the number of ML methods that are integrated
into the Bonsai ecosystem by us and by other method developers.

### 2021-11-17 application/userEngagement.tex

Bonsai's 10th anniversary will take place in 2022. The proposed machine learning package will thus leverage planned venues, events, and infrastructure to disseminate
awareness throughout the community, including presentations at conferences,
workshops, and training sessions, as well as electronic newsletters, forums and
social media articles.

The proposed development will facilitate interfacing with other programming languages
such as Python, R, and MATLAB. This will therefore attract a whole new
community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
we plan to organize hackathons to accelerate integration of their existing data
analysis and machine learning algorithms and methods into Bonsai using the new
machine learning package.

 Given the modular and integrative nature of the machine learning package, we
 plan to advertise its development early in the project, to allow interested
 partners to contribute to key
 components of the platform, such as the Python and MATLAB integrations. We
 also want to promote the co-development of algorithms by early adopters of the
 package, in true collaborative fashion. Given its modular nature and built-in
 package manager capable of handling dependency resolution and curation of
 community contributed content, Bonsai is well positioned for such an approach.

>>>>

Bonsai's 10th anniversary will take place in 2022. The proposed ML package will thus leverage planned venues, events, and infrastructure to disseminate
awareness throughout the community, including presentations at conferences,
workshops, and training sessions, as well as electronic newsletters, forums and
social media articles.

The proposed development will facilitate interfacing with other programming languages
such as Python, R, and MATLAB. This will therefore attract a whole new
community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
we plan to organize hackathons to accelerate integration of their existing data
analysis and ML algorithms and methods into Bonsai using the new
machine learning package.

 Given the modular and integrative nature of the ML package, we
 plan to advertise its development early in the project, to allow interested
 partners to contribute to key
 components of the platform, such as the Python and MATLAB integrations. We
 also want to promote the co-development of algorithms by early adopters of the
 package, in true collaborative fashion. Given its modular nature and built-in
 package manager capable of handling dependency resolution and curation of
 community contributed content, Bonsai is well positioned for such an approach.
